Everthing you wanted to know about PanDA but were truly truly afraid to ask

### 2017-01-31 PandaEvolution.tex
Reorganizing structure of the paper and starting to fix latex

\section{Evolution of PanDA WMS}
\subsection{Brief Introduction to Titan}
While PanDa  currently uses  more than 100,000  cores at well over 100 Grid sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will  require more resources  than Grid computing can possibly provide. The Worldwide LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned analysis and data processing, but it will be insufficient  for Monte Carlo (MC) production and any extra activities. Additional computing and storage  resources are therefore required.  To alleviate  these challenges, ATLAS is engaged  in an ambitious  program to expand  the current computing model to include  additional  resources such as the opportunistic   use of supercomputers   as well as commercial and academic clouds. In turn this activity drives evolution of the PanDA WMS.

@@

The most suitable task for HPCs in HEP is event generation. Event generators are mostly computational,  stand-alone code, with few input requirements.  There is no need  to stage-in much data. The event generation  in ATLAS corresponds  to
10-15\% of all jobs on the Grid. A more difficult workload to adapt  for HPCs is detector  simulation. Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.  However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.

@@

The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US. Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
Integration with Titan is the current focus for PanDA developers. The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface \cite{SAGA}. This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world. The pilot can communicate with the PanDA server from the front-end machine. The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job. ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design. Time allocation on an HPC is competitive and large projects are often preferred. About 10\% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources. The worker nodes sit idle because there are not enough of them or they do not have enough time to handle a large scale computing job. On a machine of the scale of Titan, these 10\% correspond to estimated 300M core hours per year. A system that can occupy those temporarily free nodes with smaller scale tasks would be very valuable. This offers a great possibility for PanDA to harvest the opportunistic resources on Titan and use a similar mechanism on other HPCs. Functionality has been added to the PanDA Pilot to collect information about available unused worker nodes on Titan in real time. This allows PanDA to define precisely size and duration of jobs submitted to Titan according to available free resources. First test of the system were quite successful and show great promise in increasing the resource utilization on Titan.
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
\caption{Schematic view of the PanDA system\label{fig:interface}}
\end{center}
\end{figure}

>>>>

\subsection{Brief Introduction to Titan}

While PanDa  currently uses  more than 100,000  cores at well over 100 Grid
sites with a  peak performance  of 0.3 petaFLOPS,  next LHC data taking run will
require more resources  than Grid computing can possibly provide. The Worldwide
LHC Computing Grid (WLCG) infrastructure will be sufficient for the planned
analysis and data processing, but it will be insufficient  for Monte Carlo (MC)
production and any extra activities. Additional computing and storage  resources
are therefore required.  To alleviate  these challenges, ATLAS is engaged  in an
ambitious  program to expand  the current computing model to include  additional
resources such as the opportunistic   use of supercomputers   as well as
commercial and academic clouds. In turn this activity drives evolution of the
PanDA WMS\@.

@@

The most suitable task for HPCs in HEP is event generation. Event generators are
mostly computational,  stand-alone code, with few input requirements.  There is
no need  to stage-in much data. The event generation  in ATLAS corresponds  to
10--15\% of all jobs on the Grid. A more difficult workload to adapt  for HPCs
is detector  simulation. Simulation tasks tend to be tied to the framework of a
particular simulation and often require  database access for geometry and
relevant run conditions.  However,  there is a  lot to be gained  since
simulation  tasks correspond to 2/3 of the Grid capacity.

@@

The Titan supercomputer, current number two (number one from November 2012 until
June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing
Facility (OLCF) in Oak Ridge National Laboratory, US\@. Titan, was the first
large-scale system to use  a hybrid architecture that utilizes worker nodes with
both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.

Integration with Titan is the current focus for PanDA developers. The project
aims to integrate Titan with the PanDA system using an updated PanDA Pilot that
runs on the front-end node and submits ATLAS payloads to the worker nodes using
the local batch system (PBS) via the SAGA (Simple API for Grid Applications)
interface~\cite{SAGA}. This solves several complications of  running on HPC
worker nodes,  including the lack of connectivity to outside world. The pilot
can communicate with the PanDA server from the front-end machine. The
interactive front-end  machines and the worker nodes  use  a shared  file system
which makes  it  easy  for the pilot to stage-in  any input files that are
required by the payload and stage-out the produced output files after the
payload has finished at the end of the job. ATLAS Tier 1 computing  center at
Brookhaven National Lab is currently  used for data stage out from Titan but in
principle  that can be any Grid site.

\paragraph*{HPC Backfill} HPC facilities are geared towards  large scale jobs by
design. Time allocation on an HPC is competitive and large projects are often
preferred. About 10\% of capacity on a  typical HPC machine  is unused  due to
mismatches between job sizes and available  resources. The worker nodes sit idle
because there are not enough of them or they do not have enough time to handle a
large scale computing job. On a machine of the scale of Titan, these 10\%
correspond to estimated 300M core hours per year. A system that can occupy those
temporarily free nodes with smaller scale tasks would be very valuable. This
offers a great possibility for PanDA to harvest the opportunistic resources on
Titan and use a similar mechanism on other HPCs. Functionality has been added to
the PanDA Pilot to collect information about available unused worker nodes on
Titan in real time. This allows PanDA to define precisely size and duration of
jobs submitted to Titan according to available free resources. First test of the
system were quite successful and show great promise in increasing the resource
utilization on Titan.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
\caption{Schematic view of the PanDA system\label{fig:interface}}
\end{center}
\end{figure}

### 2017-01-31 bckground.tex
Reorganizing structure of the paper and starting to fix latex

\subsection{Workflows and Workloads}

@@

\subsection{The ATLAS Monte Carlo Workflow}

The distinction between an abstract and concrete workflow instances highlights
an important element of the execution process of a workflow. Abstract workflows
can be interpreted as a formal description of the requirements for the workflow
execution, while concrete workflows can be interpreted as a formal description
of the workflow's execution process. The process of resource selection matches
the resource capabilities to the workflow requirements making possible the
transition from execution requirements to execution process.

Workload managers are those components of workflow systems or, more in general,
of distributed applications that enable resource selection, acquisition, and
use. The set of capabilities required by a workload manager mostly depends on
the heterogeneity and dynamicity of both the workflow and the resources.
Heterogeneity is a measure of the diversity of the workflow requirements and
resource capabilities that need to be matched to enable the execution of the
workflow's tasks; dynamicity is a measure of how much these requirements and
capabilities varies during execution.

The PanDA workload management system was initially designed to manage the
execution of static, relatively homogeneous tasks on dynamic but homogeneous
resources. Progressively, PanDA has evolved to enable the management of more
heterogeneous tasks and resources. In this paper we describe how the design and
architecture of PanDA has enabled executions of multi-threaded tasks on Titan,
currently the largest high performance computing (HPC) resource available in the
USA for scientific research.

Traditionally, the ATLAS workflow has been based on single-core tasks, executed
on Grid resources to process large amount (2TB) of input data divided into
discrete I/O units called ``events''. This approach has many merits: (i)
simplified process of resource selection under the assumption of homogeneous
resources and execution environments; (ii) opportunistic distribution of compute
tasks across resources based on contingent availability; (iii) relative
robustness due to the ability of re-executing failed tasks; and (iv)
multi-tiered distribution and storage of events' data.

Despite all the advantages, ATLAS workflows have evolved to include
compute-intense tasks that can benefit from multithreaded execution on HPC
machines. The Monte Carlo workflow~\cite{} is one of the largest workflows
executed by the ATLAS experiment and at least one of its stages can benefit from
large amount of parallelism. The Geant4 toolkit is used to simulate the passage
of particles through matter and\ldots

\begin{figure}
  \includegraphics[width=\columnwidth]{figures/atlas_workflow.pdf}
  \caption{ATLAS Monte Carlo workflow.}
\label{fig:atlas_workflow}
\end{figure}

>>>>

\subsection{Terminology}

\mtnote{To be moved to the end of the introduction.}

@@

### 2017-01-31 overall.tex
Reorganizing structure of the paper and starting to fix latex

\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
\label{sec:panda_titan}

@@

When ATLAS was originally deployed at the OLCF, all shared libraries were
placed at the Spider 2 Lustre file system. However, it was quickly discovered
that the I/O pattern at the Athena stage was detrimental to the overall Spider
2 file system performance. Since Spider 2 is a center-wide file system and
shared by all OLCF resources and users, this resulted in lower overall
interactive and metadata performance at OLCF. As can be seen in
Listing~\ref{mdstrace} the metadata I/O activity for ATLAS exhibits a spike
corresponding to the beginning of the runs (i.e., the Athena stage) before
tapering off.

@@

After moving the ATLAS release on Lustre to the NFS-exported read-only
directories, the metadata contention problem has resolved and a two orders of
magnitude metadata performance improvement (i.e., ~6,300 seconds on Lustre vs.
~1,500 seconds on NFS) has been observed for ATLAS.  Also, for both cases the
input file is placed on memory disk on a Titan compute nodes and for Lustre the
validation step took on average 1,320 seconds, while it was 40 seconds on NFS,
showing a three orders of magnitude performance improvement.

@@

To better understand the I/O impact of ATLAS PanDA project on Titan
supercomputing environment we analyzed 1,175 jobs ran on the week of
10/25/2016, for a total of 174 hours. Table~\ref{panda-olcf-stats} shows the
overall statistical breakdown of the observed file I/O impact of ATLAS at OLCF.
Figures \ref{fig:atlas-titan-io-read} and \ref{fig:atlas-titan-io-written} show
the file read and write I/O histograms for these 1,175 jobs, respectively.
Figures \ref{fig:atlas-titan-file-open} and \ref{fig:atlas-titan-file-close}
show the file open() and close() metadata load histograms of the same 1,175
ATLAS jobs, respectively.

@@

ATLAS jobs are file read I/O heavy, as can be seen in
Table~\ref{panda-olcf-stats}. The maximum amount read by any ATLAS job in
aggregate in this observed period was less than 250 GB and the maximum amount
written in aggregate was less than 75 GB. The average amount read per job is 20
GB and average amount written is 6 GB.

@@

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_file_open_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file open() histogram on Titan for week of 10/25/16.}
    \label{fig:atlas-titan-file-open}
\end{figure}

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_file_close_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file close() histogram on Titan for week of 10/25/16.}
    \label{fig:atlas-titan-file-close}
\end{figure}

On the metadata I/O breakdown, ATLAS PanDA jobs yield 23 file open() operations
and 5 file close() operations per second. As can be seen from figures
\ref{fig:atlas-titan-file-open} and \ref{fig:atlas-titan-file-close}, they
exhibit similar distributions. The maximum number of file open() operations are
around 170 on average and the maximum number of file close() operations are 39
on average per second per job. The total number of file open() operations is
172,089,760 for the observed window of 174 hours, while the total number of
file close() operations stand at 40,132,992 for the total 1,175 ATLAS PanDA
jobs in the same observation window. The difference between these two values is
puzzling and it is under investigation at the time of writing this paper.
{\sarp{can someone take a look at this? I would expect at least similar values
for number of opens and closes and there is vast difference. Can it be that
ATLAS jobs do a lot of opens without following closes, or perhaps multiple
opens for the same files over and over?}}

\begin{table*}[t]
\centering
\begin{tabular}{lllllllll}
 & Num. Nodes & Duration (s) & Read (GB) & Written (GB) & GB Read/nodes & GB Written/nodes & open() & close() \\
Min & 1 & 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\
Max & 300 & 7,452 & 241.06 & 71.71 & 0.81670 & 0.23903 & 1,260,185 & 294,908 \\
Average & 35.66 & 6,280.82 & 20.36 & 6.87 & 0.38354 & 0.16794 & 146,459.37 & 34,155.74 \\
Std. Dev. & 55.33 & 520.99 & 43.90 & 12.33 & 0.19379 & 0.03376 & 231,346.55 & 53,799.08
\end{tabular}
\caption{The Statistical breakdown of the I/O impact of 1,175 PanDA jobs executed at OLCF for the week of 10/25/16}
\label{panda-olcf-stats}
\end{table*}

>>>>

@@

When ATLAS was originally deployed at the OLCF, all shared libraries were
placed at the Spider 2 Lustre file system. However, it was quickly discovered
that the I/O pattern at the Athena stage was detrimental to the overall Spider
2 file system performance. Since Spider 2 is a center-wide file system and
shared by all OLCF resources and users, this resulted in lower overall
interactive and metadata performance at OLCF\@. As can be seen in
Listing~\ref{mdstrace} the metadata I/O activity for ATLAS exhibits a spike
corresponding to the beginning of the runs (i.e., the Athena stage) before
tapering off.

@@

After moving the ATLAS release on Lustre to the NFS-exported read-only
directories, the metadata contention problem has resolved and a two orders of
magnitude metadata performance improvement (i.e., \~6,300 seconds on Lustre vs.\
\~1,500 seconds on NFS) has been observed for ATLAS\@.  Also, for both cases the
input file is placed on memory disk on a Titan compute nodes and for Lustre the
validation step took on average 1,320 seconds, while it was 40 seconds on NFS,
showing a three orders of magnitude performance improvement.

@@

To better understand the I/O impact of ATLAS PanDA project on Titan
supercomputing environment we analyzed 1,175 jobs ran on the week of 10/25/2016,
for a total of 174 hours. Table~\ref{panda-olcf-stats} shows the overall
statistical breakdown of the observed file I/O impact of ATLAS at OLCF\@.
Figures~\ref{fig:atlas-titan-io-read} and~\ref{fig:atlas-titan-io-written} show
the file read and write I/O histograms for these 1,175 jobs, respectively.
Figures~\ref{fig:atlas-titan-file-open} and~\ref{fig:atlas-titan-file-close}
show the file $open()$ and $close()$ metadata load histograms of the same 1,175
ATLAS jobs, respectively.

@@

ATLAS jobs are file read I/O heavy, as can be seen in
Table~\ref{panda-olcf-stats}. The maximum amount read by any ATLAS job in
aggregate in this observed period was less than 250 GB and the maximum amount
written in aggregate was less than 75 GB\@. The average amount read per job is
20 GB and average amount written is 6 GB\@.

@@

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_file_open_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file $open()$ histogram on Titan for week of 10/25/16.}
\label{fig:atlas-titan-file-open}
\end{figure}

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_file_close_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file $close()$ histogram on Titan for week of 10/25/16.}
\label{fig:atlas-titan-file-close}
\end{figure}

On the metadata I/O breakdown, ATLAS PanDA jobs yield 23 file $open()$
operations and 5 file $close()$ operations per second. As can be seen from
figures~\ref{fig:atlas-titan-file-open} and~\ref{fig:atlas-titan-file-close},
they exhibit similar distributions. The maximum number of file $open()$
operations are around 170 on average and the maximum number of file $close()$
operations are 39 on average per second per job. The total number of file
$open()$ operations is 172,089,760 for the observed window of 174 hours, while
the total number of file $close()$ operations stand at 40,132,992 for the total
1,175 ATLAS PanDA jobs in the same observation window. The difference between
these two values is puzzling and it is under investigation at the time of
writing this paper. {\sarp{can someone take a look at this? I would expect at
least similar values for number of opens and closes and there is vast
difference. Can it be that ATLAS jobs do a lot of opens without following
closes, or perhaps multiple opens for the same files over and over?}}

\begin{table*}[t]
\centering
\begin{tabular}{lllllllll}
 & Num. Nodes & Duration (s) & Read (GB) & Written (GB) & GB Read/nodes & GB Written/nodes & $open()$ & $close()$ \\
Min & 1 & 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\
Max & 300 & 7,452 & 241.06 & 71.71 & 0.81670 & 0.23903 & 1,260,185 & 294,908 \\
Average & 35.66 & 6,280.82 & 20.36 & 6.87 & 0.38354 & 0.16794 & 146,459.37 & 34,155.74 \\
Std. Dev. & 55.33 & 520.99 & 43.90 & 12.33 & 0.19379 & 0.03376 & 231,346.55 & 53,799.08
\end{tabular}
\caption{The Statistical breakdown of the I/O impact of 1,175 PanDA jobs executed at OLCF for the week of 10/25/16}
\label{panda-olcf-stats}
\end{table*}

### 2017-01-31 paper.tex
initial drop on atlas mds performance improvement
minor re-structuring.
finishing the write up for the I/O impact section
Preliminary description of Figure 2
Reorganizing structure of the paper and starting to fix latex

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[10pt, conference, compsocconf]{IEEEtran}
% Add the compsocconf option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

@@

@@

\newcommand{\sarp}[1]{{\color{green}\textit{sarp: {#1}}}}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Everthing you wanted to know about PanDA but were truly truly afraid to ask}

% author names and affiliations
% use a multiple column layout for up to two different
% affiliations

\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
line 2: name of organization, acronyms acceptable\\
line 3: City, Country\\
line 4: Email: name@xyz.com}
\and
\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
line 2: name of organization, acronyms acceptable\\
line 3: City, Country\\
line 4: Email: name@xyz.com}
}

@@

\begin{abstract}
 Experiments at the Large Hadron Collider (LHC) face unprecedented   computing   challenges.  Heterogeneous  resources are distributed worldwide, thousands of physicists analyzing the data need remote access to hundreds of computing  sites, the volume of processed data is beyond  the exabyte  scale, and data processing requires more than billions of hours of computing usage per year. The PanDA (Production and Distributed Analysis) system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment. In the process, the old batch job paradigm of computing in HEP was discarded in favor of a far more flexible and scalable model. The success of PanDA at the LHC is leading to widespread adoption and testing by other experiments.  PanDA is  the first exascale  workload management   system  in HEP,  already operating  at a million computing jobs per day, and processing over an exabyte of data in
2013. We will describe the design and implementation  of PanDA, present  data on the performance  of PanDA at the LHC,  and discuss plans  for future  evolution of the system  to meet new challenges of scale, heterogeneity  and increasing user base.
\end{abstract}

@@

% ----------------------------------------------------------------------------
% I
% ----------------------------------------------------------------------------

\input{introduction}
--------------------------------------------------------------------
\begin{enumerate}
  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
  \item Why are WMS needed?
  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
  \item \ldots
  \item Overview/Summary of paper.
\end{enumerate}

% ----------------------------------------------------------------------------
% II
% ----------------------------------------------------------------------------

\input{Background}
\begin{enumerate}
  \item Brief overview of distributed CI.
  \item Workload vs Workflow. (DONE)
  \item Distinguish between WFMS and WLMS.
  \begin{enumerate}
    \item Role of Workload MS: help ``hide'' and federate heterogeneous resources.
  \end{enumerate}
\end{enumerate}

--------------------------------------------------------------------

% ----------------------------------------------------------------------------
% III
% ----------------------------------------------------------------------------

\input{WMSLiterature}
--------------------------------------------------------------------
\begin{enumerate}
  \item Other workload management systems (WMS) (not exclusive)
  \begin{enumerate}
    \item glidein-WMS
    \item Dirac
    \item ALICE-AlieN
  \end{enumerate}
{enumerate}
  \item Comparison and contrast
  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
\end{enumerate}

% ----------------------------------------------------------------------------
% IV
% ----------------------------------------------------------------------------
\section{PanDA Overview}\label{sec:panda_overview}
--------------------------------------------------------------------
\begin{enumerate}
  \item PanDA WMS
  \begin{enumerate}
    \item State Diagrams
    \item Control Flow Diagrams
    \item Software Module and Architecture diagrams.
  \end{enumerate}
  \item PANDA Pilot
\end{enumerate}
--------------------------------------------------------------------

@@

@@

% ----------------------------------------------------------------------------
% V
% ----------------------------------------------------------------------------

\section{Deploying PanDA on a Leadership-scale system}
\label{sec:panda_deployment}

\begin{enumerate}
  \item Challenges
  \item Opportunities
  \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
  \item Take-aways, lessons learned
\end{enumerate}

% ----------------------------------------------------------------------------
% VI
% ----------------------------------------------------------------------------
\input{perf}

%\section{Performance Characterization}
%\label{sec:panda_performance}
%
%\begin{enumerate}
%  \item Internal PanDA performance
%  \item External PanDA performance (workload specific on underlying system/platform)
%\end{enumerate}

% ----------------------------------------------------------------------------
% VII
% ----------------------------------------------------------------------------
\input{overall}

@@

% ----------------------------------------------------------------------------
% VIII
% ----------------------------------------------------------------------------

\section{PANDA: Future/RoadMap}
\input{Experiments}
\label{sec:panda_roadmap}

 \section{Conclusion}\label{sec:conclusion}
The PanDA  system was developed to meet the scale and complexity of LHC distributed computing for the ATLAS experiment.  In the process,  the old batch job paradigm  of computing in HEP was  discarded  in favor of a  far more flexible and scalable  model. The success  of PanDA  at the LHC is leading to widespread adoption and testing by other experiments. PanDA  is the first exascale  workload management system in HEP, already operating at a million computing jobs per day, and processing over an exabyte of data in 2013. Next LHC run will pose massive computing  challenges. With a  doubling of the beam  energy  and luminosity as  well as an increased  need for  simulates  data, the data volume is expected to increase with a factor 5-6 or more. Storing and processing  this amount of data is a  challenge   that cannot be resolved with the currently existing  computing  resources in ATLAS. To resolve this challenge, ATLAS is turning to commercial  as well as academic Cloud services and HPCs via the PanDA system. Also the work underway is enabling the use of PanDA by new scientific collaborations and communities as a means  of leveraging  extreme scale computing  resources with a low barrier of entry. The technology base provided by the PanDA system will enhance the usage of a variety  of high-performance computing resources available to basic research.

\input{ack}

@@

\bibliographystyle{plain}
\bibliography{bibliography}

>>>>

\documentclass[10pt, conference, compsocconf]{IEEEtran}

@@

\usepackage{listings}
\usepackage{comment}

@@

% -----------------------------------------------------------------------------
% BEGIN DOCUMENT
% -----------------------------------------------------------------------------
\begin{document}

\title{Everthing you wanted to know about PanDA but were truly truly afraid to
ask}

% author names and affiliations
% use a multiple column layout for up to two different
% affiliations
\author{\IEEEauthorblockN{Authors Name/s per 1st Affiliation (Author)}
\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
line 2: name of organization, acronyms acceptable\\
line 3: City, Country\\
line 4: Email: name@xyz.com}
\and
\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
line 2: name of organization, acronyms acceptable\\
line 3: City, Country\\
line 4: Email: name@xyz.com}
}

@@

% ----------------------------------------------------------------------------
% ABSTRACT
% ----------------------------------------------------------------------------
\begin{abstract}
Experiments at the Large Hadron Collider (LHC) face unprecedented computing
challenges. Heterogeneous resources are distributed worldwide, thousands of
physicists analyzing the data need remote access to hundreds of computing
sites, the volume of processed data is beyond the exabyte scale, and data
processing requires more than billions of hours of computing usage per year. The
PanDA (Production and Distributed Analysis) system was developed to meet the
scale and complexity of LHC distributed computing for the ATLAS experiment. In
the process, the old batch job paradigm of computing in HEP was discarded in
favor of a far more flexible and scalable model. The success of PanDA at the LHC
is leading to widespread adoption and testing by other experiments. PanDA is
the first exascale workload management system in HEP, already operating at
a million computing jobs per day, and processing over an exabyte of data in
2013. We will describe the design and implementation of PanDA, present data on
the performance of PanDA at the LHC, and discuss plans for future evolution
of the system to meet new challenges of scale, heterogeneity and increasing
user base.
\end{abstract}

@@

% ----------------------------------------------------------------------------
% I - INTRODUCTION
% ----------------------------------------------------------------------------
\input{introduction}

\begin{enumerate}
  \item Problem that has been solved: executing many applications on
  heterogeneous distributed resources.
  \item Why are WMS needed?
  \item Increasing importance of WMS w.r.t complex and scalable applications on
  heterogeneous distributed resources.
  \item \ldots
  \item Overview/Summary of paper.
\end{enumerate}

% ----------------------------------------------------------------------------
% II - BACKGROUND AND RELATED WORK
% ----------------------------------------------------------------------------
\section{Background and Related Work}
\label{sec:bkgrd_literature}

\begin{enumerate}
  \item Other workload management systems (WMS) (not exclusive)
  \begin{enumerate}
    \item glidein-WMS
    \item Dirac
    \item ALICE-AlieN
  \end{enumerate}
  \begin{enumerate}
    \item Comparison and contrast
    \item Core features of a general WMS, i.e., minimal complete model of a WMS.
  \end{enumerate}
\end{enumerate}

\input{WMSLiterature}

\begin{enumerate}
  \item Brief overview of distributed CI.
  \item Workload vs Workflow. (DONE)
  \item Distinguish between WFMS and WLMS.
  \begin{enumerate}
    \item Role of Workload MS: help ``hide'' and federate heterogeneous resources.
  \end{enumerate}
\end{enumerate}

% Please do not correct bckground to background.
\input{bckground}

% ----------------------------------------------------------------------------
% III - PanDA OVERVIEW
% ----------------------------------------------------------------------------
\section{PanDA Overview}
\label{sec:panda_overview}

\begin{enumerate}
  \item PanDA WMS
  \begin{enumerate}
    \item State Diagrams
    \item Control Flow Diagrams
    \item Software Module and Architecture diagrams.
  \end{enumerate}
  \item PANDA Pilot
\end{enumerate}

% ----------------------------------------------------------------------------
% IV - ATLAS MONTE CARLO WORKFLOW
% ----------------------------------------------------------------------------
\section{ATLAS Monte Carlo Workflow}
\label{sec:mc_workflow}

\input{workflow}

% ----------------------------------------------------------------------------
% V - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
% ----------------------------------------------------------------------------
\section{Deploying PanDA on a Leadership-scale system}
\label{sec:panda_deployment}

\begin{enumerate}
  \item Why? To support a specific stage of the ATLAS Monte Carlo Workflow.
  \item Challenges
  \item Opportunities
  \item Highlight distinction of PanDA on wide-area distributed systems vs on TITAN
  \item Take-aways, lessons learned
\end{enumerate}

@@

% ----------------------------------------------------------------------------
% VI - EVOLUTION OF PANDA WMS
% ----------------------------------------------------------------------------
% \section{Evolution of PanDA WMS}
% \label{sec:panda_evolution}

@@

% ----------------------------------------------------------------------------
% VI - PERFORMANCE CHARACTERIZATION
% ----------------------------------------------------------------------------
% \section{Performance Characterization}
% \label{sec:panda_performance}
%
% \input{perf}

% ----------------------------------------------------------------------------
% VI - PUTTING IT ALL TOGETHER: EXPERIENCE OF PANDA ON TITAN
% ----------------------------------------------------------------------------
\section{Putting it all together: Experience of PanDA on TITAN} % (fold)
\label{sec:panda_titan}

\input{overall}

@@

% ----------------------------------------------------------------------------
% VII - THE NEXT GENERATION EXECUTOR
% ----------------------------------------------------------------------------
% \section{PANDA: Future/RoadMap}
\section{PANDA: The Next Generation Executor}
\label{sec:panda_roadmap}

\begin{enumerate}
    \item Rationale and design (why)
    \item Architecture (how)
    \item Integration
    \item Characterization (experiments)
\end{enumerate}

\input{Experiments}

% ----------------------------------------------------------------------------
% VIII - CONCLUSIONS
% ----------------------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}

The PanDA  system was developed to meet the scale and complexity of LHC
distributed computing for the ATLAS experiment.  In the process,  the old batch
job paradigm  of computing in HEP was  discarded  in favor of a  far more
flexible and scalable  model. The success  of PanDA  at the LHC is leading to
widespread adoption and testing by other experiments. PanDA  is the first
exascale  workload management system in HEP, already operating at a million
computing jobs per day, and processing over an exabyte of data in 2013. Next LHC
run will pose massive computing  challenges. With a  doubling of the beam
energy  and luminosity as  well as an increased  need for  simulates  data, the
data volume is expected to increase with a factor 5-6 or more. Storing and
processing  this amount of data is a  challenge   that cannot be resolved with
the currently existing  computing  resources in ATLAS. To resolve this
challenge, ATLAS is turning to commercial  as well as academic Cloud services
and HPCs via the PanDA system. Also the work underway is enabling the use of
PanDA by new scientific collaborations and communities as a means  of leveraging
extreme scale computing  resources with a low barrier of entry. The technology
base provided by the PanDA system will enhance the usage of a variety  of
high-performance computing resources available to basic research.

% ----------------------------------------------------------------------------
% ACKNOWLEDGEMENTS
% ----------------------------------------------------------------------------
\input{ack}

@@

% ----------------------------------------------------------------------------
% REFERENCES
% ----------------------------------------------------------------------------
\bibliographystyle{plain}
\bibliography{bibliography}
